---
title: "MARONE: STMO PROJECT"
---

# Data Pre-processing

## Data extraction & First Insight

```{r}
# Load the data
data <- read.csv("insurance.csv")

# Display the structure of the dataframe
str(data)

# Display the first few rows of the dataframe
head(data)
```

## Shape and data type

```{r}
# Check the shape of the dataframe
data_shape <- dim(data)

# Display the result
cat("Number of rows:", data_shape[1], "\n")
cat("Number of columns:", data_shape[2], "\n")
```

```{r}
# Check the data types of the features
feature_types <- sapply(data, class)

# Display the result
print(feature_types)
```

## Converting the character into categorical feature and also children

```{r}
# Convert non-numerical features to categorical features
data$sex <- as.factor(data$sex)
data$smoker <- as.factor(data$smoker)
data$region <- as.factor(data$region)
data$children <- as.factor(data$children)

# Check the data types after conversion
feature_types_after_conversion <- sapply(data, class)
print(feature_types_after_conversion)
```

## Checking Missing Values

```{r}
# Check for missing values in the entire dataframe
any_missing <- any(is.na(data))

# Check for missing values in each column
missing_per_column <- colSums(is.na(data))

# Display the results
cat("Any missing values in the dataframe:", any_missing, "\n")
cat("Missing values per column:\n")
print(missing_per_column)
```

```{r}
library(visdat)
vis_dat(data)
```

# Exploratory and Descriptive Data Analysis

## Summary and distribution of numerical features

### Summary function

```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)

# Custom function to calculate mode
calculate_mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

# Assuming 'data' is a global variable or defined somewhere accessible

# Function to summarize and plot a feature
summary_plot <- function(feature_name, bin_width) {
  
  # Extract the specified feature
  feature <- data[[feature_name]]
  
  # Summary statistics
  n <- length(feature)
  mean_val <- mean(feature)
  var_val <- var(feature)
  q1_val <- quantile(feature, 0.25)
  q3_val <- quantile(feature, 0.75)
  mode_val <- calculate_mode(feature)
  max_val <- max(feature)
  min_val <- min(feature)
  
  # Confidence Intervals
  # Confidence interval for the mean
  mean_ci <- t.test(x = feature)$conf.int
  
  # Confidence interval for the variance
  alpha <- 0.05  # significance level
  df <- n - 1  # degrees of freedom
  upper <- df * var_val / qchisq(alpha / 2, df)
  lower <- df * var_val / qchisq(1 - alpha / 2, df)
  var_ci <- c(lower, upper)
  
  custom_summary <- data.frame(
    Mean = mean_val,
    Mean_CI_Lower = mean_ci[1],
    Mean_CI_Upper = mean_ci[2],
    Variance = var_val,
    Variance_CI_Lower = var_ci[1],
    Variance_CI_Upper = var_ci[2],
    Q1 = q1_val,
    Q3 = q3_val,
    Mode = mode_val,
    Max = max_val,
    Min = min_val
  )
  
  print(custom_summary)


  
  # Plots
  histogram_plot <- ggplot(data, aes(x = feature)) +
    geom_histogram(aes(y = ..density..), binwidth = bin_width, fill = "lightblue", color = "black", alpha = 0.7) +
    geom_density(color = "red", linetype = "dashed") +  # Add density curve
    labs(title = paste("Histogram of", feature_name), x = feature_name, y = "Density") +
    theme_minimal()

  
  violin_plot <- ggplot(data, aes(x = factor(1), y = feature)) +
    geom_violin(fill = "lightgreen", color = "black", alpha = 0.7) +
    labs(title = paste("Violin Plot of", feature_name), x = "", y = feature_name) +
    theme_minimal() +
    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
  
  boxplot_plot <- ggplot(data, aes(x = factor(1), y = feature)) +
    geom_boxplot(fill = "lightcoral", color = "black", alpha = 0.7) +
    labs(title = paste("Boxplot of", feature_name), x = "", y = feature_name) +
    theme_minimal() +
    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
  
  cumulative_plot <- ggplot(data, aes(x = feature)) +
    stat_ecdf(geom = "step", color = "blue") +
    labs(title = paste("Cumulative Frequency Distribution of", feature_name), x = feature_name, y = "Cumulative Frequency") +
    theme_minimal()
  
  # Arrange the plots in one figure
  grid.arrange(histogram_plot, violin_plot, boxplot_plot, cumulative_plot, ncol = 2)
}
```

### Summary and distribution of : Age

```{r}
summary_plot("age", 5)
```

### Summary and distribution of : BMI

```{r}
summary_plot("bmi", 5)
```

### Summary and distribution of : Charges

```{r}
summary_plot("charges", 1700)
```

## Summary and distribution categorical features

### Functions that plot

```{r}
summary_categorical_feature <- function(feature, alpha = 0.05) {
  # Plotting
  plot <- ggplot(data, aes(x = get(feature), fill = get(feature))) +
    geom_bar(color = "black", stat = "count") +
    labs(title = paste("Count Plot of", feature),
         x = feature, y = "Count") +
    scale_fill_manual(values = c("red", "blue", "green", "magenta", "yellow", "brown")) +  # Adjust the colors as needed
    theme_minimal()
  
  print(plot)
  
  # Compute proportions and confidence intervals
  feature_data <- data[[feature]]
  prop_table <- table(feature_data)
  total_obs <- length(feature_data)
  result_df <- data.frame(Category = character(), Proportion = numeric(), Lower_CI = numeric(), Upper_CI = numeric(), stringsAsFactors = FALSE)
  
  for (category in names(prop_table)) {
    count_category <- prop_table[[category]]
    proportion <- count_category / total_obs
    conf_interval <- prop.test(count_category, total_obs, conf.level = 1 - alpha)$conf.int
    result_row <- data.frame(Category = category,
                             Proportion = proportion,
                             Lower_CI = conf_interval[1],
                             Upper_CI = conf_interval[2],
                             stringsAsFactors = FALSE)
    result_df <- rbind(result_df, result_row)
  }
  
  return(result_df)
}
```

```{r}
# Function to create a count plot of one categorical feature grouped by another
count_plot_two_categorical <- function(feature1, feature2) {
  ggplot(data, aes(x = get(feature1), fill = get(feature2))) +
    geom_bar(position = "dodge", color = "black", stat = "count") +
    labs(title = paste("                     Count Plot of", feature1, "Grouped by", feature2),
         x = feature1, y = "Count") +
    theme_minimal() +
    theme(legend.position = "top")
}
```

```{r}
library(ggplot2)

# Function to create histograms, violin plots, and box plots
plot_numerical_by_categorical <- function(numerical_feature, categorical_feature, bin_width) {
  
  # Histogram
  histogram_plot <- ggplot(data, aes(x = get(numerical_feature), fill = get(categorical_feature))) +
    geom_histogram(binwidth = bin_width, color = "black", alpha = 0.7, position = "identity") +
    labs(title = paste("Superposed Histogram of", numerical_feature, "by", categorical_feature),
         x = numerical_feature, y = "Count") +
    theme_minimal()
  
  # Violin plot
  violin_plot <- ggplot(data, aes(x = get(categorical_feature), y = get(numerical_feature), fill = get(categorical_feature))) +
    geom_violin(color = "black", alpha = 0.7) +
    labs(title = paste("Violin Plot of", numerical_feature, "by", categorical_feature),
         x = categorical_feature, y = numerical_feature) +
    theme_minimal()
  
  # Box plot
  box_plot <- ggplot(data, aes(x = get(categorical_feature), y = get(numerical_feature), fill = get(categorical_feature))) +
    geom_boxplot(color = "black", alpha = 0.7) +
    labs(title = paste("Box Plot of", numerical_feature, "by", categorical_feature),
         x = categorical_feature, y = numerical_feature) +
    theme_minimal()
  
  #  Arrange the plots in one figure
  grid.arrange(histogram_plot, violin_plot, box_plot, ncol = 2)
}
```

```{r}
# load the package rstatix
library(rstatix)

# define the function
test_distribution <- function(categorical_feature, numerical_feature) {
  # perform Kruskal-Wallis test
  result <- data %>%
    group_by(data[[categorical_feature]]) %>%
    kruskal_test(data[[numerical_feature]] ~ data[[categorical_feature]])
  
  # return the p-value
  return(result)
}
```

### The Distribution of the sex

```{r}
# Considering the entire data set
summary_categorical_feature("sex")

# Considering the smokers
count_plot_two_categorical("smoker", "sex")

# Considering the Region
count_plot_two_categorical("region", "sex")

# Considering the Children
count_plot_two_categorical("children", "sex")

# distribution of the age among the smokers
plot_numerical_by_categorical("age", "sex", 3)

# Distribution of the BMI among smokers
plot_numerical_by_categorical("bmi", "sex", 3)

# Distribution of the charges among smokers
plot_numerical_by_categorical("charges", "sex", 1300)
```

```{r}
p_value <- test_distribution("sex", "charges")
print(p_value)

p_value <- test_distribution("sex", "bmi")
print(p_value)

p_value <- test_distribution("sex", "age")
print(p_value)
```

### The Distribution of the smokers

```{r}
# Considering the entire data set
summary_categorical_feature("smoker")

# Considering the region
count_plot_two_categorical("region", "smoker")

# Considering the number of children
count_plot_two_categorical("children", "smoker")

# distribution of the age among the smokers
plot_numerical_by_categorical("age", "smoker", 3)

# Distribution of the BMI among smokers
plot_numerical_by_categorical("bmi", "smoker", 3)

# Distribution of the charges among smokers
plot_numerical_by_categorical("charges", "smoker", 1300)
```

```{r}
p_value <- test_distribution("smoker", "charges")
print(p_value)

p_value <- test_distribution("smoker", "bmi")
print(p_value)

p_value <- test_distribution("smoker", "age")
print(p_value)
```

### The Distribution of the children

```{r}
# Considering the entire data set
summary_categorical_feature("children")

# Considering the region
count_plot_two_categorical("region", "children")

# distribution of the age among the smokers
plot_numerical_by_categorical("age", "children", 3)

# Distribution of the BMI among smokers
plot_numerical_by_categorical("bmi", "children", 3)

# Distribution of the charges among smokers
plot_numerical_by_categorical("charges", "children", 3500)
```

```{r}
# Assuming data is your data frame
data$many_children <- ifelse(data$children %in% c(0, 1, 2), 0, 1)
data$many_children <- as.factor(data$many_children)

# Considering the entire data set
summary_categorical_feature("many_children")

# Distribution of the charges among smokers
plot_numerical_by_categorical("charges", "many_children", 3500)

count_plot_two_categorical("smoker", "many_children")
```

```{r}
p_value <- test_distribution("many_children", "charges")
print(p_value)

p_value <- test_distribution("many_children", "bmi")
print(p_value)

p_value <- test_distribution("many_children", "age")
print(p_value)
```

```{r}
# Considering the entire data set
summary_categorical_feature("region")

# distribution of the age among the smokers
plot_numerical_by_categorical("age", "region", 3)

# Distribution of the BMI among smokers
plot_numerical_by_categorical("bmi", "region", 3)

# Distribution of the charges among smokers
plot_numerical_by_categorical("charges", "region", 3500)
```

```{r}
p_value <- test_distribution("region", "charges")
print(p_value)
```

## Variable selection

```{r}
data_new <- data[, -which(names(data) == "sex")]
data_new <- data_new[, -which(names(data_new) == "children")]
data_new <- data_new[, -which(names(data_new) == "region")]
```

## Scatter plots

```{r}
# load the package : ggplot2 & GGally
library(ggplot2)
library(GGally)

# Assuming data_new is your dataframe
ggpairs(data_new, mapping = ggplot2::aes(color = smoker), progress = FALSE)
```

## Correlation matrix

```{r}
# Select numeric variables for the correlation matrix
numeric_data <- data_new[sapply(data_new, is.numeric)]

# Create a correlation matrix plot
ggcorr(numeric_data, label = TRUE, label_alpha = TRUE, label_color = "black", hjust = 0.5, vjust = 1, palette = "RdYlBu")
```

# Hypothesis test

## Conditions verification

### function

```{r}
library(car)
library(lmtest)

# define the function
check_assumptions <- function(model) {
  # initialize a data frame to store the results
  results <- data.frame(test = character(), p_value = numeric())
  
  # check for multivariate normality
  shapiro_test <- shapiro.test(residuals(model))
  results <- rbind(results, data.frame(test = "Multivariate normality", p_value = shapiro_test$p.value))
  
  # check for multicollinearity
  vif_values <- vif(model)
  if (any(vif_values > 5)) {
    results <- rbind(results, data.frame(test = "Multicollinearity", p_value = NA))
  }
  
  # check for autocorrelation
  durbin_watson_test <- durbinWatsonTest(model)
  results <- rbind(results, data.frame(test = "Autocorrelation", p_value = durbin_watson_test$p))
  
  # check for homoscedasticity
  bp_test <- bptest(model)
  results <- rbind(results, data.frame(test = "Homoscedasticity", p_value = bp_test$p.value))
  
  # return the results
  return(results)
}
```

### Variables : smoker, bmi, age, many_children

```{r}
# fit a linear regression model
lm_model <- lm(charges ~ ., data = data_new)

# use the function
results <- check_assumptions(lm_model)
print(results)

summary(lm_model)
```

### Variables : smoker, bmi, age

```{r}
# Remove the "many_children" column
data_rmv_many_children <- data_new[, -which(names(data_new) == "many_children")]

# fit a linear regression model
lm_model_rmv_many_children <- lm(charges ~ ., data = data_rmv_many_children)

# use the function
results <- check_assumptions(lm_model_rmv_many_children)
print(results)

summary(lm_model_rmv_many_children)
```

### Variables : bmi, age

```{r}
# Remove the "smoker" and "many_children" column
data_rmv_manY_children_smoker <- data_new[, -which(names(data_new) == "smoker")]
data_rmv_manY_children_smoker <- data_rmv_manY_children_smoker[, -which(names(data_rmv_manY_children_smoker) == "many_children")]

# fit a linear regression model
lm_model_rmv_manY_children_smoker <- lm(charges ~ ., data = data_rmv_manY_children_smoker)

# use the function
results <- check_assumptions(lm_model_rmv_manY_children_smoker)
print(results)

summary(lm_model_rmv_manY_children_smoker)
```

### Variables : bmi, age (smokers and non-smokers separated)

```{r}
# Take the smokers
data_smoker_many_child <- data_new[data_new$smoker == "yes", ]

# Remove the "smoker" and "many_children" column
data_smoker_many_child <- data_smoker_many_child[, -which(names(data_smoker_many_child) == "smoker")]
data_smoker_many_child <- data_smoker_many_child[, -which(names(data_smoker_many_child) == "many_children")]

# fit a linear regression model
lm_model_smoker_many_child <- lm(charges ~ ., data = data_smoker_many_child)

# use the function
results <- check_assumptions(lm_model_smoker_many_child)
print(results)

summary(lm_model_smoker_many_child)



# Take the smokers
data_smoker_no_many_child <- data_new[data_new$smoker == "no", ]

# Remove the "smoker" and "many_children" column
data_smoker_no_many_child <- data_smoker_no_many_child[, -which(names(data_smoker_no_many_child) == "smoker")]
data_smoker_no_many_child <- data_smoker_no_many_child[, -which(names(data_smoker_no_many_child) == "many_children")]

# fit a linear regression model
lm_model_smoker_no_many_child <- lm(charges ~ ., data = data_smoker_no_many_child)

# use the function
results <- check_assumptions(lm_model_smoker_no_many_child)
print(results)

summary(lm_model_smoker_no_many_child)
```

### Variables : smoker, bmi, age (many_children/few_children) for non-smokers

```{r}
# Take the smokers
data_smoker_many_child <- data_new[data_new$smoker == "no", ]
data_smoker_many_child <- data_smoker_many_child[data_smoker_many_child$many_children == 1, ]

# Remove the "smoker" and "many_children" column
data_smoker_many_child <- data_smoker_many_child[, -which(names(data_smoker_many_child) == "smoker")]
data_smoker_many_child <- data_smoker_many_child[, -which(names(data_smoker_many_child) == "many_children")]

# fit a linear regression model
lm_model_smoker_many_child <- lm(charges ~ ., data = data_smoker_many_child)

# use the function
results <- check_assumptions(lm_model_smoker_many_child)
print(results)

summary(lm_model_smoker_many_child)




# Take the smokers
data_smoker_no_many_child <- data_new[data_new$smoker == "no", ]
data_smoker_no_many_child <- data_smoker_no_many_child[data_smoker_no_many_child$many_children == 1, ]

# Remove the "smoker" and "many_children" column
data_smoker_no_many_child <- data_smoker_no_many_child[, -which(names(data_smoker_no_many_child) == "smoker")]
data_smoker_no_many_child <- data_smoker_no_many_child[, -which(names(data_smoker_no_many_child) == "many_children")]

# fit a linear regression model
lm_model_smoker_no_many_child <- lm(charges ~ ., data = data_smoker_no_many_child)

# use the function
results <- check_assumptions(lm_model_smoker_no_many_child)
print(results)

summary(lm_model_smoker_no_many_child)
```

### Variables : smoker, bmi, age (many_children/few_children) for smokers

```{r}
# Take the smokers
data_smoker <- data_new[data_new$smoker == "yes", ]

# Remove the "smoker" and "many_children" column
data_smoker <- data_smoker[, -which(names(data_smoker) == "smoker")]
data_smoker <- data_smoker[, -which(names(data_smoker) == "many_children")]

# fit a linear regression model
lm_model_smoker <- lm(charges ~ ., data = data_smoker)

# use the function
results <- check_assumptions(lm_model_smoker)
print(results)

summary(lm_model_smoker)


# Take the smokers
data_non_smoker <- data_new[data_new$smoker == "no",]

# Remove the "smoker" and "many_children" column
data_non_smoker <- data_non_smoker[, -which(names(data_non_smoker) == "smoker")]
data_non_smoker <- data_non_smoker[, -which(names(data_non_smoker) == "many_children")]

# fit a linear regression model
lm_model_non_smoker <- lm(charges ~ ., data = data_non_smoker)

# use the function
results <- check_assumptions(lm_model_non_smoker)
print(results)

summary(lm_model_non_smoker)
```

# Model Evaluation & Prediction

## Evaluation

```{r}
library(caret)

# Set a seed for reproducibility
set.seed(123)

# Split the data into training and testing sets (80% training, 20% testing)
index <- createDataPartition(data_rmv_many_children$charges, p = 0.8, list = FALSE)
train_data <- data_rmv_many_children[index, ]
test_data <- data_rmv_many_children[-index, ]

# Fit a linear regression model on the training data
lm_model_rmv_many_children <- lm(charges ~ ., data = train_data)

# Predict on the test set
predictions <- predict(lm_model_rmv_many_children, newdata = test_data)

# Evaluate the model
metrics <- data.frame(
  RMSE = sqrt(mean((test_data$charges - predictions)^2)),
  MAE = mean(abs(test_data$charges - predictions)),
  R_squared = 1 - sum((test_data$charges - predictions)^2) / sum((test_data$charges - mean(test_data$charges))^2)
)

# Print the metrics
print(metrics)
```

## Prediction

```{r}
# Randomly select an index from the test set
random_index <- sample(1:nrow(test_data), 1)

# Extract the selected observation
selected_observation <- test_data[random_index, ]

# Extract the predictor variables for the selected observation
selected_predictors <- selected_observation[, -which(names(selected_observation) == "charges")]

# Predict the charges for the selected observation
prediction <- predict(lm_model_rmv_many_children, newdata = selected_predictors)

# Compute the prediction interval
prediction_interval <- predict(lm_model_rmv_many_children, newdata = selected_predictors, interval = "prediction")

# Print the actual value, prediction, and prediction interval
cat("Actual Value:", selected_observation$charges, "\n")
cat("Prediction:", prediction, "\n")
cat("Prediction Interval: [", prediction_interval[2], ";", prediction_interval[3], "]\n")
```

## 
